\propertitle

### 2019-12-20 maxNt_v6.tex
update

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-12-20T20:13:38+0100
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Inferring the total activity of a large neuronal population\\ from a small sample}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Inferences about large neuronal populations}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

\newcommand*{\sh}{\mathrm{\varEta}}
\newcommand*{\nat}{\textrm{nat}}
\newcommand*{\hart}{\textrm{Hart}}

@@

As mentioned in the Introduction, in the neurosciences the maximum-entropy
method has also been used as a way of quantifying the
\enquote{cooperativity} \citep[\eg][]{gersteinetal1985} or
\enquote{interaction}
\citep[\eg][]{martignonetal1995,schneidmanetal2006,shlensetal2006} or
\enquote{synchrony} \citep[\eg]{bohteetal2000,amarietal2003} of neuronal
activity. Most, if not all, such applications use the method \emph{at the
  sample level}, aiming to find a distribution for the sample without
considering that it comes from a larger population. In this section we
discuss how our proposed application bears on this kind of quantification
and compares with the application at the sample level. We also wish to
highlight some methodological caveats.% For concreteness we use an
% example with specific statistics: first and second moments (equivalent to
% mean and correlations), versus first-to-fourth moments. But our
% discussion holds for any sets of statistics.

@@

As a first step we need to translate vague terms like
\enquote{cooperativity}, \enquote{interaction}, and similar into a more
precise, operational \citep[\cf][\chap~I]{bridgman1927_r1958} notion. Here
we propose the notion of \emph{informational sufficiency}, because
\begin{enumerate*}[label=(\roman*)]
\item it seems closely related to those terms as used in the
  literature; \item it is mathematically connected with maximum-entropy
  distributions; \item it reveals important consequences of sampling.
\end{enumerate*}
Its meaning and the three points above will now be illustrated. For an
extensive discussion of this notion, probably first suggested by Kullback
\amp\ Leibler \parencite*{kullbacketal1951}, see Bernardo \etal\
\parencite*[\sect~4.5]{bernardoetal1994} and Jaynes \parencite*[\chap~8 and
\sect~14.2]{jaynes1994_r2003}.

Our belief distribution about the frequencies of the neuronal activities,
or about the activity in a new time bin, is in principle conditional on all
experimental data and statistics we have. This distribution, however, can
be such that it remains unchanged if we discard part of the data or
statistics -- for example, the third- and higher-order empirical moments. % --
% but it does change if we discard more.
This means that the discarded statistics are \emph{informationally
  irrelevant}: knowing them does not change our beliefs. The remaining
statistics -- in the example, first and second empirical moments -- are
\emph{informationally sufficient}.%  The qualification \enquote{minimally}
% will be omitted for brevity.
% This informational quasi-sufficiency is interesting because
% it may give clues about population organization and dynamics.

Now let us turn the informational sufficiency of a subset of statistics and
the informational sufficiency of another subset into two hypotheses. We can
then ask: which hypothesis is more probable, given the data? equivalently,
what is the ratio of their probabilities? By translating our question about
\enquote{cooperativity} into a hypothesis-testing problem about sufficiency,
we can find a well-defined answer determined by the
probability-calculus. % The ratio of
% the two probabilities gives us a natural measure of the
% For example, do
% the data say that the first two empirical moments are probably enough to
% determine our beliefs? or is it more probable that the first four are
% needed instead?

But such a translation does more: through a mathematical theorem \citep[and
references therein]{andersen1970}[this theorem was first proved for
continuous probability distributions
by][]{koopman1936,pitman1936,darmois1935}[see][\sect~4.5.3 for further
references]{bernardoetal1994}, it leads to the automatic appearance of
maximum-entropy distributions: \emph{if a probability distribution for a
  repetitive phenomenon has some sufficient statistics, then it is a
  mixture of distributions of maximum-entropy form having those statistics
  as constraints, and vice versa}. 
% Thus, maximum-entropy distributions automatically appear (as
% approximations) from the probability-calculus when we test hypotheses
% about informational sufficiency.
Its most important consequence is that, by Bayes's formula, the probability
that some statistics are informationally sufficient can be approximated by
finding the maximum-entropy distribution constrained by those statistics
and then calculating the probability for the whole data under that
distribution. We shall shortly illustrate this kind of calculation for the
first data set.

@@

The theorem above makes it very natural to reinterpret studies about
cooperativity, interaction, and similar, based on maximum-entropy, as
studies about informational sufficiency. An important fact must then be
taken into account: \emph{informational sufficiency is not preserved under
  sampling}. If a probability distribution has some sufficient statistics,
then its sample marginals \emph{cannot} have the same sufficient
statistics, and vice versa \citep[\sect~3.1]{portamanaetal2015}; except for
trivial cases such as uniform probability
distributions. % This generally holds also among
% different marginals: if a statistics is sufficient for some marginal, then
% it will not be sufficient for another marginal.
This impossibility is known in another guise in statistical mechanics: if a
system is described by a Gibbs state, its subsystems cannot be perfectly
described by Gibbs states \citep[\eg][and references
therein]{maesetal1999}. % Mathematically this impossibility comes from the
% Pitman-Koopman theorem mentioned above, and translates into the general
% impossibility of solving a system of independent equations with more
% equations than unknowns .

This fact affects the interpretation of the results obtained from
maximum-entropy distributions. For example, if means and pairwise correlations
seem informationally sufficient for a particular sample from a brain area,
then they are probably not sufficient for the larger population of neurons
constituting that area, and vice versa. If we are interested in the
cooperativity of the neurons in a brain area, it is therefore unreliable to
use a maximum-entropy distribution constructed only for a sample
thereof. % \citep[contrast this with some literature, \eg][]{martignonetal1995,bohteetal2000,schneidmanetal2006,shlensetal2006,barreiroetal2010,ganmoretal2011,granotatedgietal2013}
The model presented here avoids this problem because the maximum-entropy
method is applied directly to the larger population, not to the sample
alone.

@@

We compare the informational sufficiency between of a set of moments, say
$M''\defd\set{1,\dotsc,m''}$, with respect to another, say
$M'\defd \set{1,\dotsc,m'}$, by introducing the quantity $\ydi(M'',M')$
defined as follows:
\begin{enumerate}[label=(\roman*)]
\item From each maximum-entropy distribution $\px(\yA \| M,N)$ for the
  larger population, built from each set of constraints $M=M',M''$,
  calculate the marginal distribution for the sample by
  \eqn~\eqref{eq:conditional_hypergeometric_t}:
\begin{equation}
  \label{eq:sample_marginal}
  \pxx(\ya \| M, N) = \sum_{\yA}\yG_{\ya \yA}\px(\yA \| M,N),
  \qquad M=M',M''.
\end{equation}
\item Calculate the relative entropies of the measured frequency
  distribution $\yf$ with respect to each sample marginal from the previous
  step, and multiply them by the number of time bins $T$:
  \begin{equation}
    \label{eq:relentropy_T}
    T\,\sh[\yf; \pxx(\ya \| M,N)] \defd
    T  \sum_{\ya} \yff_{\ya} \log\frac{\yff_{\ya}}{\pxx(\ya \| M,N)},
    \qquad M=M',M''.
  \end{equation}
\item Take the difference: 
\begin{equation}
  \label{eq:measure_suffic}
  \begin{split}
  \ydi_{N}(M'',M') &\defd
  T \,\sh[\yf; \pxx(\ya \| M',N)]
  -
  T \,\sh[\yf; \pxx(\ya \| M'',N)]
  \\ &\equiv
T  \sum_{\ya} \yff_{\ya} \log\frac{
  \sum_{\yA}\yG_{\ya\yA}\px(\yA \| M'',N)}{
  \sum_{\yA}\yG_{\ya\yA}\px(\yA \| M',N)}.
\end{split}
\end{equation}
\end{enumerate}
The quantity $\ydi_{N}(M'',M')$ so defined is positive if $M''$ is more
probable to be informationally sufficient than $M'$, and negative
otherwise.

@@

% NUMERICAL RESULTS
%
% relentropies of sample2mom, sample4mom, sample5mom, marg2mom, marg4mom, marg5mom
% -322.72,-42.8005,-39.4917,-85.7626,-4.23725,-4.20059 nats =
% -1401.56,-185.88,-171.51,-372.462,-18.4021,-18.2429 dB
%
% difference in relentropies between 4 mom and 2 mom:
% sample level: 279.92 nats = 1215.68 dB
% marg level: 81.525 nats =  354.06 dB
% difference in relentropies between 5 mom and 4 mom:
% sample level: 3.30882 nats = 14.37 dB
% marg level: 0.0366623 nats = 0.159223 dB
% correspond to prob ratios of
% sample: 3.69524\times 10^{121}, 27.3529
% marg: 2.5469\times 10^{35}, 1.03734
% 
% comparison N vs n:
% 2 mom: 236.958 nats = 1029.09 dB, 8.11758\times 10^{102}
% 4 mom: 38.5632 nats = 167.478 dB, 5.59496*10^16
% 5 mom: 35.2911 nats = 153.267 dB, 2.12185*10^15
%
% comparison_ME_sample_full_levels.nb
%

@@

Let us illustrate the ideas, the remarks, and the quantity defined so far
with our first data set. Consider for example three sets of constraints
$M_{2}$, $M_{4}$, $M_{5}$, consisting of the first two, four, five moments;
and a larger population of size $N=10\,000$. Figure~\ref{fig:plots_2_4mom}
shows the resulting distributions for the two data sets.
\begin{figure}[!p]
\centering
\subcaption[]{\label{fig:plots_2_4mom_a}}%  
\includegraphics[width=\linewidth]{_scripts/newpaper2_stensola_threemom_distributions10000_priorc.pdf}%
\\[2em]%
\subcaption[]{\label{fig:plots_2_4mom_b}}%  
\includegraphics[width=\linewidth]{_scripts/newpaper2_riehle_threemom_distributions10000_prior2.pdf}%
\\[2em]%
\caption{Maximum-entropy distributions for the normalized total activity of
  the larger population, assuming two different sets of constraining moments
  and a larger population size $N=10\,000$.}
\label{fig:plots_2_4mom}
\end{figure}
% maxent_solutions_different_moments_prior2_6moments.nb
There is a clear visual difference between the distributions based on two
(\textcolor{myred}{dashed red}) and on four (\textcolor{myblue}{solid
  blue}) moments; for the first data set, panel
\subcaptionref{fig:plots_2_4mom_a}, they even differ in the number of
modes. On the other hand there is barely any visual difference between the
distributions based on four and on five (\textcolor{myyellow}{dotted
  yellow}) moments. From these visual differences we immediately conclude
that two moments are less likely to be informationally sufficient than
four; but four should almost as likely as five. We shall see that the
measure~\eqref{eq:measure_suffic} reflects what we immediately see.

Compare these plots with those obtained applying the model at the sample
level (equivalent to $N=n$). With the $M_{2}$ and $M_{4}$ constraints we
obtain the distributions partly shown in
\fig~\ref{fig:comparisons_marginals} as \textcolor{myred}{red empty
  circles} for two moments, and \textcolor{myblue}{blue empty squares} for
four moments. The measured frequencies (black triangles) are also shown.
The two distributions can only be distinguished on a logarithmic scale, and
only in their tails. It would be difficult to visually assess which is more
likely to be informationally sufficient.

@@

Calculation of the weight of evidence~\eqref{eq:measure_suffic} for the
first data set and $N=10\,000$, corresponding to
\fig~\ref{fig:plots_2_4mom_a}, leads to the following results:
\begin{equation}
  \label{eq:diff_suff_N}
  \begin{split}
  \ydi_{N}(M_{4},M_{2}) &= 81\;\nat = 35\;\hart,
\\    \ydi_{N}(M_{5},M_{4}) &= 0.037\;\nat = 0.016\;\hart,
  \end{split}
\end{equation}
where the Hartley ($\hart$) denotes base-10 logarithms
\citep[\sect~C.4]{iso2009}[called \enquote{ban} and used by Turing and Good
in their code-breaking work at Bletchley Park:
see][]{good1985%,good1950,good1969
}[\sect~4.2]{jaynes1994_r2003}. In words, our data are 35 orders of
magnitude more probable by assuming sufficiency of $M_{4}$ than by assuming
sufficiency of $M_{2}$. But the data are about as probable
($10^{0.016} = 1.04$) by assuming sufficiency of $M_{4}$ as of $M_{5}$. The
qualitative visual information of \fig~\ref{fig:plots_2_4mom_a} was
therefore reliable.

For the model at the sample level, that is, with $N=n$, the weights of
evidence are
\begin{equation}
  \label{eq:diff_suff_n}
  \begin{split}
  \ydi_{n}(M_{4},M_{2}) &= 280\;\nat = 1\,220\;\hart,
\\    \ydi_{n}(M_{5},M_{4}) &= 3.3\;\nat = 1.4\;\hart.
\end{split}
\end{equation}
So the application at the sample level says that the data are 1\,220 orders
of magnitude more probable conditional on four moments than conditional on
two, and $10^{1.4}=25$ times more probable conditional on five moments than
on four moments. These results are different from the ones obtained from
the larger population, and they would be hard to deduce visually from a
look at the distributions of \fig~\ref{fig:comparisons_marginals}.

@@

It is in fact possible to use the measure~\eqref{eq:measure_suffic} and its
probabilistic interpretation~\eqref{eq:bayes_factor} to compare the
probability of the data $\yf$ conditional on the hypothesis $(M_{4},N)$ of
four-moment sufficiency at the larger-population level, $N=10\,000$, versus
the hypothesis $(M_{4}, n)$ of four-moment sufficiency at the sample level,
$N=n=65$. We obtain
\begin{equation}
  \label{eq:diff_suff_n}
  \ydi[(M_{4}, N), (M_{4}, n)] = 39\;\nat = 17\;\hart.
\end{equation}
% that is, the data are 17 orders of magnitude more probable under the first
% hypothesis than under the second. 
This result is also shown in
\fig~\ref{fig:comparisons_marginals}. We see that the four-moment
distribution constructed at the larger-population level
(\textcolor{myblue}{blue filled squares}) fits the measured frequency
distribution (black triangles) more closely than the corresponding
sample-level distribution (\textcolor{myblue}{blue empty squares}).

@@

The calculations and conclusions presented in the preceding sections depend
on the size $N$ of the larger population. The larger population can't be
the whole brain, of course. How large should or can $N$ be in our formulae?

The crucial point is our belief distribution for the activity of the sample
\emph{conditional} on the activity of the larger population. It is given by
the hypergeometric distribution~\eqref{eq:conditional_hypergeometric},
reprinted here:
\begin{equation}
  \labelbis{eq:conditional_hypergeometric}
  \yG_{\ya\yA} \defd \pf(\ya \|\yA, n,N)=
    \binom{\yA}{\ya}\binom{N-\yA}{n-\ya}\binom{N}{n}^{-1}
\equiv \binom{n}{\ya}\binom{N-n}{\yA-\ya}\binom{N}{\yA}^{-1}.
\end{equation}
This distribution leads to the equality of factorial
moments~\eqref{eq:equal_m_factorialmoments}, on which our model rests. This
conditional probability, characteristic of \enquote{drawing without
  replacement}
\citep[\eg][\chap~3]{jaynes1994_r2003}, % [\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}
ensues when our degree of belief that the $i$th unit (ball or neuron) will
be drawn has the same value for all $i \in \set{1,\dotsc,N}$. Thus,
consider the pool of all neurons which we believe -- with equal degree for
each neuron -- could have been recorded. $N$ is the size of that pool. $N$
therefore depends on factors such as: the shape, dimensions, and technical
specifications of the recording probe; the inaccuracy in the insertion of
the probe, leading for example to slightly different insertion points or
angles; the density of neurons around the probe. But it also depends on the
homogeneity of the brain region where the recording was made: if we believe
that it doesn't matter whether the probe had been inserted in some other
point of the same brain region, then
formula~\eqref{eq:conditional_hypergeometric} is appropriate. $N$ can
therefore only be assessed case by case.

>>>>

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-12-21T00:40:05+0100
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Inferring the total activity of a large neuronal population\\ from a small sample}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Inferences about large neuronal populations}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

\newcommand*{\sh}{\mathrm{\varEta}}
\newcommand*{\nat}{\textrm{nat}}
\newcommand*{\bit}{\textrm{bit}}
\newcommand*{\hart}{\textrm{Hart}}

@@

As mentioned in the Introduction, in the neurosciences the maximum-entropy
method has also been used to quantify the \enquote{cooperativity}
\citep[\eg][]{gersteinetal1985} or \enquote{interaction}
\citep[\eg][]{martignonetal1995,schneidmanetal2006,shlensetal2006} or
\enquote{synchrony} \citep[\eg]{bohteetal2000,amarietal2003} of neuronal
activity. Most, if not all, such applications use the method \emph{at the
  sample level}, that is, they find a distribution for the sample without
considering that it comes from a larger population. In this section we
discuss how our proposed application to a larger population bears on this
kind of quantification and compares with the application at the sample
level. %We also highlight some methodological caveats.% For concreteness we use an
% example with specific statistics: first and second moments (equivalent to
% mean and correlations), versus first-to-fourth moments. But our
% discussion holds for any sets of statistics.

@@

As a first step, terms like \enquote{cooperativity}, \enquote{interaction},
and similar need to be translated into a more precise, operational
\citep[\cf][\chap~I]{bridgman1927_r1958} notion. Here we propose the notion
of \emph{informational sufficiency}, because
\begin{enumerate*}[label=(\roman*)]
\item it seems closely related to those terms as used in the
  literature; \item it is mathematically connected with maximum-entropy
  distributions; \item it reveals important consequences of sampling.
\end{enumerate*}
Its meaning will now be briefly explained. For an extensive discussion of
this notion, which was probably first suggested by Kullback \amp\ Leibler
\parencite*{kullbacketal1951}, see Bernardo \etal\
\parencite*[\sect~4.5]{bernardoetal1994} and Jaynes \parencite*[\chap~8 and
\sect~14.2]{jaynes1994_r2003}.

Our belief distribution about the frequencies of the neuronal activities,
or about the activity in a new time bin, is conditional on all experimental
data and statistics we have collected. This distribution, however, can be such that
it remains unchanged if we discard part of the data or statistics -- for
example, the third- and higher-order empirical moments. % --
% but it does change if we discard more.
This means that the discarded statistics are \emph{informationally
  irrelevant}: knowing them does not change our beliefs. The remaining
statistics -- in the example, first and second empirical moments -- are
\emph{informationally sufficient}.%  The qualification \enquote{minimally}
% will be omitted for brevity.
% This informational quasi-sufficiency is interesting because
% it may give clues about population organization and dynamics.

Now let us turn the informational sufficiencies of two different subsets of
statistics into two hypotheses. We can then ask: which hypothesis is more
probable, given the data? equivalently, what is the ratio of their
probabilities? By translating our question about \enquote{cooperativity}
into a hypothesis-testing problem about sufficiency, we can find a
well-defined answer determined by the probability-calculus. % The ratio of
% the two probabilities gives us a natural measure of the
% For example, do
% the data say that the first two empirical moments are probably enough to
% determine our beliefs? or is it more probable that the first four are
% needed instead?

But such a translation does more: through a mathematical theorem \citep[and
references therein]{andersen1970}[this theorem was first proved for
continuous probability distributions
by][]{koopman1936,pitman1936,darmois1935}[see][\sect~4.5.3 for further
references]{bernardoetal1994}, it leads to the automatic appearance of
maximum-entropy distributions: \emph{if a probability distribution for a
  repetitive phenomenon has some sufficient statistics, then it is a
  mixture of distributions of maximum-entropy form having those statistics
  as constraints, and vice versa}. 
% Thus, maximum-entropy distributions automatically appear (as
% approximations) from the probability-calculus when we test hypotheses
% about informational sufficiency.
The most important consequence of this theorem is that, by Bayes's formula,
the probability that some statistics are informationally sufficient can
approximately be found by first building the maximum-entropy distribution
constrained by those statistics, and then calculating the probability for
the whole data under that distribution. We shall shortly illustrate this
procedure with the data sets of \sect~\ref{sec:application}.

@@

Owing to the theorem above, studies about cooperativity (and similar ideas)
that use maximum-entropy can be naturally reinterpreted as studies about
informational sufficiency. An important fact must then be taken into
account: \emph{informational sufficiency is not preserved under sampling}.
If a probability distribution has some sufficient statistics, then its
sample marginals \emph{cannot} have the same sufficient statistics, and
vice versa \citep[\sect~3.1]{portamanaetal2015}; except for trivial cases
(\eg\ uniform probability distributions). % This generally holds also among
% different marginals: if a statistics is sufficient for some marginal, then
% it will not be sufficient for another marginal.
This impossibility is known in another guise in statistical mechanics: if a
system is described by a Gibbs state, its subsystems cannot be perfectly
described by Gibbs states \citep[\eg][and references
therein]{maesetal1999}. % Mathematically this impossibility comes from the
% Pitman-Koopman theorem mentioned above, and translates into the general
% impossibility of solving a system of independent equations with more
% equations than unknowns .

This fact affects the interpretation of the results obtained from
maximum-entropy distributions. For example, if means and pairwise
correlations seem informationally sufficient for a sample from a brain
area, then they are probably not sufficient for the larger population of
neurons constituting that area, and vice versa. If we are interested in the
cooperativity of the neurons in a brain area, it is therefore unreliable to
use a maximum-entropy distribution constructed only for a sample
thereof. % \citep[contrast this with some literature, \eg][]{martignonetal1995,bohteetal2000,schneidmanetal2006,shlensetal2006,barreiroetal2010,ganmoretal2011,granotatedgietal2013}
The model presented here avoids this problem because it targets the larger
population, not the sample alone.

@@

We can compare the informational sufficiency between of a set of moments, say
$M''\defd\set{1,\dotsc,m''}$, with respect to another, say
$M'\defd \set{1,\dotsc,m'}$, by introducing the quantity $\ydi(M'',M')$
defined as follows:
\begin{enumerate}[label=(\roman*)]
\item From each maximum-entropy distribution $\px(\yA \| M,N)$ for the
  larger population, \eqn~\eqref{eq:app_maxent_pop_first}, built from each
  set of constraints $M=M',M''$, calculate the marginal distribution for
  the sample by \eqn~\eqref{eq:conditional_hypergeometric_t}:
\begin{equation}
  \label{eq:sample_marginal}
  \pxx(\ya \| M, N) = \sum_{\yA}\yG_{\ya \yA}\px(\yA \| M,N),
  \qquad M=M',M''.
\end{equation}
\item Calculate the relative entropies of the measured frequency
  distribution $\yf$ with respect to each sample marginal from the previous
  step, and multiply them by the number of time bins $T$:
  \begin{equation}
    \label{eq:relentropy_T}
    T\,\sh[\yf; \pxx(\ya \| M,N)] \defd
    T  \sum_{\ya} \yff_{\ya} \log\frac{\yff_{\ya}}{\pxx(\ya \| M,N)},
    \qquad M=M',M''.
  \end{equation}
\item Take the difference: 
\begin{equation}
  \label{eq:measure_suffic}
  \begin{split}
  \ydi_{N}(M'',M') &\defd
  T \,\sh[\yf; \pxx(\ya \| M',N)]
  -
  T \,\sh[\yf; \pxx(\ya \| M'',N)]
  \\ &\equiv
T  \sum_{\ya} \yff_{\ya} \log\frac{
  \sum_{\yA}\yG_{\ya\yA}\px(\yA \| M'',N)}{
  \sum_{\yA}\yG_{\ya\yA}\px(\yA \| M',N)}.
\end{split}
\end{equation}
\end{enumerate}
The quantity $\ydi_{N}(M'',M')$ so defined is positive if $M''$ is more
probable to be informationally sufficient than $M'$, and negative
otherwise.

@@

@@

Let us illustrate the quantity just defined and the previous discussion
about sufficiency and sampling by using our data sets. Consider for example
three sets of constraints $M_{2}$, $M_{4}$, $M_{5}$, consisting of the
first two, four, five moments; and a larger population of size $N=10\,000$.
Figure~\ref{fig:plots_2_4mom} shows the resulting distributions for the two
data sets.
\begin{figure}[!p]
\centering
\subcaption[]{\label{fig:plots_2_4mom_a}}%  
\includegraphics[width=\linewidth]{_scripts/newpaper2_stensola_threemom_distributions10000_priorc.pdf}%
\\[2em]%
\subcaption[]{\label{fig:plots_2_4mom_b}}%  
\includegraphics[width=\linewidth]{_scripts/newpaper2_riehle_threemom_distributions10000_prior2.pdf}%
\\[2em]%
\caption{Maximum-entropy distributions for the normalized total activity of
  the larger population, assuming two different sets of constraining moments
  and a larger population size $N=10\,000$.}
\label{fig:plots_2_4mom}
\end{figure}
% maxent_solutions_different_moments_prior2_6moments.nb
There is a clear visual difference between the distributions based on two
(\textcolor{myred}{dashed red}) and on four (\textcolor{myblue}{solid
  blue}) moments; for the first data set, panel
\subcaptionref{fig:plots_2_4mom_a}, they even differ in the number of
modes. On the other hand there is barely any visual difference between the
distributions based on four and on five (\textcolor{myyellow}{dotted
  yellow}) moments. From these visual differences we immediately conclude
that two moments are less likely to be informationally sufficient than
four; but four should be almost as likely as five. We shall see that the
measure~\eqref{eq:measure_suffic} reflects this swift visual judgement.

Compare these plots with those obtained applying the model at the sample
level (equivalent to $N=n$) to the first data set. With the $M_{2}$ and
$M_{4}$ constraints we obtain the distributions partly shown in
\fig~\ref{fig:comparisons_marginals} for two (\textcolor{myred}{red empty
  circles}) and four (\textcolor{myblue}{blue empty squares}) moments. The
measured frequencies (black triangles) are also shown. The two
distributions can only be distinguished on a logarithmic scale, and only in
their tails. It would be difficult to visually assess which is more likely
to be informationally sufficient.

@@

% NUMERICAL RESULTS
%
% relentropies of sample2mom, sample4mom, sample5mom, marg2mom, marg4mom, marg5mom
% -322.72,-42.8005,-39.4917,-85.7626,-4.23725,-4.20059 nats =
% -1401.56,-185.88,-171.51,-372.462,-18.4021,-18.2429 dB
%
% difference in relentropies between 4 mom and 2 mom:
%** sample level: 279.92 nats = 403.839 bit = 1215.68 dB
%** marg level: 81.525 nats =  117.616 bit = 354.06 dB
% difference in relentropies between 5 mom and 4 mom:
%** sample level: 3.30882 nats = 4.77362 bit = 14.37 dB
%** marg level: 0.0366623 nats = 0.0543836 bit = 0.159223 dB
% correspond to prob ratios of
% sample: 3.69524\times 10^{121}, 27.3529
% marg: 2.5469\times 10^{35}, 1.03734
% 
% comparison N vs n:
% 2 mom: 236.958 nats = 341.858 bit = 1029.09 dB, 8.11758\times 10^{102}
% 4 mom: 38.5632 nats = 55.635 bit = 167.478 dB, 5.59496*10^16
% 5 mom: 35.2911 nats = 50.9157 bit = 153.267 dB, 2.12185*10^15
%
% comparison_ME_sample_full_levels.nb
%
Calculation of the weight of evidence~\eqref{eq:measure_suffic} for the
first data set and $N=10\,000$, corresponding to
\fig~\ref{fig:plots_2_4mom_a}, leads to the following results:
\begin{equation}
  \label{eq:diff_suff_N}
  \begin{split}
  \ydi_{N}(M_{4},M_{2}) &= 118\;\bit \equiv 35\;\hart,
\\    \ydi_{N}(M_{5},M_{4}) &= 0.054\;\bit \equiv 0.016\;\hart,
  \end{split}
\end{equation}
where the Hartley ($\hart$) denotes base-10 logarithms
\citep[\sect~C.4]{iso2009}[called \enquote{ban} and used by Turing and Good
in their code-breaking work at Bletchley Park:
see][]{good1985%,good1950,good1969
}[\sect~4.2]{jaynes1994_r2003}. In words, our data are 35 orders of
magnitude more probable by assuming sufficiency of $M_{4}$ than by assuming
sufficiency of $M_{2}$. But the data are about as probable
($10^{0.016} = 1.04$) by assuming sufficiency of $M_{4}$ as of $M_{5}$. The
qualitative visual information of \fig~\ref{fig:plots_2_4mom_a} was
therefore reliable.

For the model at the sample level, that is, with $N=n$, the weights of
evidence are
\begin{equation}
  \label{eq:diff_suff_n}
  \begin{split}
  \ydi_{n}(M_{4},M_{2}) &= 400\;\bit = 120\;\hart,
\\    \ydi_{n}(M_{5},M_{4}) &= 4.8\;\bit = 1.4\;\hart.
\end{split}
\end{equation}
So the application at the sample level says that the data are 120 orders of
magnitude more probable conditional on $M_{4}$ than on $M_{2}$, and
$10^{1.4}=25$ times more probable conditional on $M_{5}$ than on $M_{4}$.
These results are different from those of the application at the larger
population, and they would be hard to deduce visually from the
distributions of \fig~\ref{fig:comparisons_marginals}.

@@

It is possible to use the measure~\eqref{eq:measure_suffic} and its
probabilistic interpretation~\eqref{eq:bayes_factor} to compare the
probability of the data $\yf$ conditional on the hypothesis $(M_{4},N)$ of
four-moment sufficiency at the larger-population level, $N=10\,000$, versus
the hypothesis $(M_{4}, n)$ of four-moment sufficiency at the sample level,
$N=n=65$. We obtain
\begin{equation}
  \label{eq:diff_suff_n}
  \ydi[(M_{4}, N), (M_{4}, n)] = 56\;\bit = 17\;\hart.
\end{equation}
% that is, the data are 17 orders of magnitude more probable under the first
% hypothesis than under the second. 
This result is also shown in
\fig~\ref{fig:comparisons_marginals}. We see that the four-moment
distribution constructed at the larger-population level
(\textcolor{myblue}{blue filled squares}) fits the measured frequency
distribution (black triangles) more closely than the corresponding
sample-level distribution (\textcolor{myblue}{blue empty squares}).

@@

The calculations and conclusions presented in the preceding sections depend
on the size $N$ of the larger population. The larger population cannot be
the whole brain, of course. How large should or can $N$ be in our formulae?

The crucial point is our belief distribution for the activity of the sample
\emph{conditional} on the activity of the larger population. It is given by
the hypergeometric distribution~\eqref{eq:conditional_hypergeometric},
reprinted here:
\begin{equation}
  \labelbis{eq:conditional_hypergeometric}
  \yG_{\ya\yA} \defd \pf(\ya \|\yA, n,N)=
    \binom{\yA}{\ya}\binom{N-\yA}{n-\ya}\binom{N}{n}^{-1}
\equiv \binom{n}{\ya}\binom{N-n}{\yA-\ya}\binom{N}{\yA}^{-1}.
\end{equation}
This distribution leads to the equality of factorial
moments~\eqref{eq:equal_m_factorialmoments}, on which our model rests. This
conditional probability, characteristic of \enquote{drawing without
  replacement}
\citep[\eg][\chap~3]{jaynes1994_r2003}, % [\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}
ensues when we believe, with the same degree, that any of the $N$ units
(balls or neurons) could be drawn or sampled. Thus, consider the pool of
all neurons which we believe -- with equal degree for each neuron -- could
have been recorded. Then $N$ is the size of that pool. $N$ therefore depends on
factors such as: the shape, dimensions, and technical specifications of the
recording probe; the inaccuracy in the insertion of the probe, leading for
example to slightly different insertion points or angles; the density of
neurons around the probe. But it also depends on the homogeneity of the
brain region where the recording was made: if we believe that it doesn't
matter whether the probe had been inserted in some other point of the same
brain region, then formula~\eqref{eq:conditional_hypergeometric} is
appropriate. $N$ can therefore only be assessed case by case.
